---
title: "Final Project - Fencing Champions"
author: "Samuel Cohen, Alex Schweizer"
date: "4/26/20"
output: html_notebook
---

# Question: Throughout the history of fencing, which countries and which athletes within those countries excelled in the sport of fencing throughout time and earned medals in international tournaments, such as the FIE World Fencing Championships, and the Olympics?

## Setup:

```{r}
library(xml2)
library(rvest)
library(ggplot2)
library(tidyr)
library(dplyr)
library(DataComputing)
library(tidyverse)
#library(maps)
```

```{r}
fie <- "https://en.wikipedia.org/wiki/World_Fencing_Championships#World_champions"
fielist <- fie %>%
  read_html() %>%
  html_nodes(css = "table") %>%
  html_table(fill = TRUE)
```
```{r}
olympics <- "https://en.wikipedia.org/wiki/List_of_Olympic_medalists_in_fencing_(men)"
olympicslist <- olympics %>%
  read_html() %>%
  html_nodes(css = "table") %>%
  html_table(fill = TRUE)
```

### Links to data sources: 

FIE Championships:    https://en.wikipedia.org/wiki/World_Fencing_Championships#World_champions

Olympics:    https://en.wikipedia.org/wiki/List_of_Olympic_medalists_in_fencing_(men)

## About the first data source: 

The first dataset contains charts of data from the FIE (Federation Internationale d'Escrime) World Fencing Championships, such as the medal count of each country, the gold medalists of each event for each year and weapon, and the top 10 international fencers in each weapon with the highest medal count. This data would be held within FIE's website, but apparently there is a Wikipedia page for it, saving me the time. The Wikipedia page was sourced by former fencer, Richard Cohen (not related), and externally linked to the FIE webpage. This data has been recorded since the first ever tournament in 1921, and is kept recorded in the archives for the history books. 

Here are some of the tables in the dataset:

```{r}
medalranks <- fielist[[6]] %>%
  filter(row_number() < 43)
medalranks
```

*Each case of this table is a ranking of the number of gold, silver, and bronze medals a country has earned.

```{r}
Sabers <- fielist[[9]]
Sabers
```

*Each case in this table is the champion individuals or Country's teams in a given year.

## About the Second data source:

The Second dataset contains charts of data of the champions of the men's fencing event of the Summer Olympics. all of this data is collected by the International Olympics Committee, or IOC for short; however, there is no provider on whom published the data onto Wikipedia. The results of each Olympic Games are collected as records of history. 

Here are some of the tables from the data source:

```{r}
Olympic_Saber <- olympicslist[[5]]
Olympic_Saber
```

Each Case in this table is the Gold, Silver, and Bronze medallists of an Olympic Games.

```{r}
Olympic_S_Team <- olympicslist[[6]]
Olympic_S_Team
```

Each case in this table are the Gold, Silver, and Bronze teams of an Olympic Games.


## Data Wrangling 

```{r}
View(Sabers)
View(OlympicMedals)
View(Olympic_S_Team)
View(medalranks)
View(Olympic_Saber)
```


- (A) General Data Wrangling: 

- Examples: filter, mutate, summarise, arrange, group_by, etc.

```{r}
Gold_Olympics <- OlympicMedals %>% select(Nation, Gold) 
Gold_FIE <- medalranks %>% select(Nation, Gold) 
Total_Olympics <- OlympicMedals %>% select(Nation, Total)
Total_FIE <- medalranks %>% select(Nation, Total)

```


- (B) Join Operation: 


```{r}
JoinMedal <- medalranks %>% select(Rank, Nation)
Joined <- JoinMedal %>% left_join(OlympicMedals %>% select(Rank, Nation), by= "Rank") 

View(Joined)
```


- (C) Spread or Gather: 


```{r}

```


- (D) Reg Ex: 


```{r}
#
Details <- Olympic_Saber %>% mutate(New= gsub("details", "", Games))

#  
Years <- Details %>% extractMatches("([0-9][0-9][0-9][0-9])", New) %>% mutate(Years= match1) %>% select(Games, Years)

#
Locations <- Details %>% mutate(Locations = gsub("[0-9][0-9][0-9][0-9]", "", New))

Joined_RegEx <- Years %>% left_join(Locations, by="Games") %>% select(Years, Locations, Gold, Silver, Bronze)

View(Joined_RegEx)


#VIEW NEW COLUMNS FOR OLYMPIC SABER

#abbreviation <- OlympicMedals %>% mutate(gsub("[(FRA)]$", " ", Nation))
#abbreviation

```


- (E) Reduction/Transformation: 


```{r}
Total_Games <- Olympic_Saber %>% summarise(total= sum(n())-1)
Total_Olympics %>% mutate(Average_Total_Medals_per_Year = Total/28)
```

(F) .Rmd Line number(s) for use of user-defined functions:



## Data Visualization

Below are carious forms of Data visualizations. The data used for this project has put a big limit on the required list. There is not many types of graphs to use this kind of information, so coming up with idea of the visualizations to use have been a bit of a challenge.

### (A) .Rmd Line number(s) for use of mulitple different geoms:  

Both plots below are plots of the same data, but different geoms were used. The first plot uses geom_jitter, and each dot is a case of the total medal count of each country, and how much of the total medals are gold medals. The second plot is uses geom_bar, and shows the number of total medals each country has won, and reorder by the number of gold medals. This data can be useful regarding the opening question because it is information on which countries has been the most dominant during the history of competitive fencing. Looking at the plots, it is no surprise to one of us that at least two or three European countries are on the top of the list.

```{r}
ggplot(data = medalranks, aes(x = Total, y = Gold)) +
  geom_jitter()
```

```{r}
ggplot(data=medalranks,aes(x=reorder(Nation,Gold),y=Total ))+geom_bar(stat='identity',position='stack', width=.9) + theme(axis.text.x=element_text(angle=60,hjust=1)) 
```

### (B) .Rmd Line number(s) for use of multiple aesthetics: 

The below plot is the Olympics Medals count of each country, with a couple of aesthetics added, such as a reorder, a fill, and a theme of the x-axis. This data is relevant to the research question for same reason the above plots are. This data is interesting when comparing this plot to the bar plot above because it is weird how the three countries with the top total medal counts are the same on both plots. 

```{r}
OlympicMedals <- olympicslist[[12]] %>%
  filter(row_number() < 34)

ggplot(data=OlympicMedals,aes(x=reorder(Nation,Total),y=Total,order=reorder(Gold,Gold),fill=Gold))+geom_bar(stat='identity',position='stack', width=.9) + theme(axis.text.x=element_text(angle=60,hjust=1)) 
```

### (D) .Rmd Line number(s) for use of leaflet maps: 

The plot below is a global choropleth map, showing the total number of medals each country has won. This data is interesting because most of the countries that have been dominant in the FIE Tournament are grouped together in the map. Russia does have a good amount of medals, and everywhere else along just looks like a void.

```{r}
WorldMap(medalranks, key = Nation, fill = Total)
```


## To Conclude:

Regarding all of the data, we have found that Italy has been the most dominant country in the FIE Tournament's history, and France the most dominant country during the Olympics' history. 

To one of us, the results may be surprising, but to the other, who is a fencer, not at all. From experience, the United States does have a strong fencing environment, but some European countries do make our competition look like childs play. 






